---
title: "Introduction to the CHL26predictor R Package"
author: "Fong Chun Chan"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Introduction
***

This vignette gives a high level overview on how to use the CHL26predictor R package to generate CHL26 scores. The model consists of 26 genes (3 housekeepers and 23 predictor genes/features)

```{r, message = FALSE}
library("knitr")
library("CHL26predictor")
library("dplyr")
library("ggplot2")
library("reshape2")
```

```{r}
kable(CHL26.model.coef.df, caption = "Overview of the Features in Model")
```

# Generating CHL26 Scores
***

To generate the CHL26 scores for the predictor, the R package simply requires an expression matrix as input. The R package provides a small test expression matrix to use with this vignette:

```{r}
CHL26.test.exprs.mat
```

These expression values have been generated from the sampling of count data from a possion distribution with the lambda values sample a uniform distribution.  We'll first need to normalize the gene expression matrix:

```{r normalizer, message = FALSE, fig.width = 7, fig.cap = "Normalizer Values"}
normalizer <- get_normalizer(CHL26.test.exprs.mat)
CHL26.test.exprs.norm.mat <- normalize_mat(CHL26.test.exprs.mat, normalizer)

normalizer %>%
  melt(value.name = "normalizer") %>%
  mutate(sampleID = rownames(.)) %>%
  arrange(normalizer) %>%
  ggplot(aes(x = factor(sampleID, levels = sampleID), y = normalizer)) +
  geom_bar(stat = "identity") +
  xlab("Sample ID") +
  ylab("Normalizer") +
	theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

Now that the gene expression matrix is normalized, we can generate the CHL26 scores:

```{r CHL26_score, message = FALSE, fig.width = 7, fig.cap = "Distribution of CHL26 Scores"}
scores.df <- get_CHL26_scores(CHL26.test.exprs.norm.mat)
scores.df <- arrange(scores.df, score)

scores.df %>%
  ggplot(aes(x = factor(sampleID, levels = sampleID), y = score)) +
  geom_bar(stat = "identity") +
  xlab("Sample ID") +
  ylab("CHL26 Score") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

# Classifying Samples into High and Low Risk
***

With the CHL26 scores, we can next classify samples into high and low risk by using the published threshold of 0.6235.

```{r CHL26_score_risk, fig.width = 7, fig.cap = "Risk Class of CHL26 Score"}
risk.thres <- 0.6235
scores.risk.df <- scores.df %>%
  mutate(riskClass = ifelse(score >= risk.thres, "High", "Low"))

scores.risk.df %>%
  ggplot(aes(x = factor(sampleID, levels = sampleID), y = score, 
             fill = riskClass)) +
  geom_bar(stat = "identity") +
  xlab("Sample ID") +
  ylab("CHL26 Score") +
  geom_hline(y = risk.thres, col = "red", linetype = "dotted") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

> As this is synthetic data, the proportion of the high to low-risk class does NOT reflect the true expected proportion.
